---
title: "SOMbrero"
output: html_document
date: '2022-05-31'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)#
#knitr::opts_knit$set(root.dir = '~/Documents/data/')
```

## Basic package description
```{r Libraries, echo=FALSE}
library(SOMbrero)
library(cluster)
library(enrichR)
library(paletteer)
```

`trainSOM()` -- standard version of the algorithm on data frame or matrix with numerical columns.
Input :
  - `x.data` : data set used to train the SOM, matrix or data frame with col : numerical variables, rows : obs
  - other arguments : see`help(initSOM)`
Output :
  - `somRes` object, see`help(trainSOM)`
  
## Dataset loading
```{r}
# The following setting is important, do not omit.
options(stringsAsFactors = FALSE)
rnaseq <- read.csv("DataRnaseq", header = T, sep = " ")

dim(rnaseq)
names(rnaseq)[1:10]
typeof(rnaseq)

X <- as.matrix(t(rnaseq))
dim(X)
#head(X)
```

`X` : RNA-seq data as a matrix with rows representing genes and columns representing samples

## Test basic SOM
We'll try a basic SOM on a subset of the RNA-seq data.
```{r}
mySOM <- trainSOM(x.data = X[1:100,], nb.save = 100, verbose = TRUE,
                  maxit = 1000)
```

`nb.save` is necessary to be able to plot the energy.  
```{r}
plot(mySOM, what = "energy")
```
### Results of the clustering

```{r}
table(mySOM$clustering)
plot(mySOM, what="obs", type="hitmap")
```
The number of prototypes (clusters) can be changed with parameter `dimension`, default is `dimension = c(5,5)`.

### Assessing the relevance of each variable 
```{r}
summary(mySOM)
```
Here, all variables have significantly different means among the different clusters and can thus be considered to be relevant for the clustering definition.

### Clustering interpretation
Mean for each patient
```{r}
plot(mySOM, what = "obs", type = "color", variable = 1)
plot(mySOM, what = "obs", type = "color", variable = 2)
plot(mySOM, what = "obs", type = "color", variable = 3)
```

Distance between prototypes :
```{r}
plot(mySOM, what = "prototypes", type = "poly.dist", show.names = FALSE)
```
### Analyze the projection quality
```{r}
quality(mySOM)
```
The topographic error value varies between 0 (good projection quality) and 1 (poor projection quality).
The quantization error is an unbounded positive number. The closer it is to 0, the better the projection quality.

### Building super classes from the resulting SOM
```{r}
plot(superClass(mySOM))
```
To create `k` super clusters :  
```{r}
mySC <- superClass(mySOM, k = 3)
summary(mySC)
```
```{r}
plot(mySC, plot.var = FALSE)
plot(mySC, type = "grid")
plot(mySC, what = "prototypes", type = "color", variable = 10)
plot(mySC, what = "prototypes", type = "poly.dist")
```

## Test different parameters
### Radius type : Gaussian VS Letremy

```{r}
gaussSOM <- trainSOM(X , dimension = c(10,10), topo = "square",
                    affectation = "standard", nb.save = 30, verbose = TRUE,
                    maxit = 1500, radius.type = "gaussian")
save(gaussSOM, file = 'gaussianSOM.Rdata')
letSOM <- trainSOM(X , dimension = c(10,10), topo = "square",
                    affectation = "standard", nb.save = 30, verbose = TRUE,
                    maxit = 1500, radius.type = "letremy")
save(letSOM, file = 'letremySOM.Rdata')
```
```{r}
Gauss.nrg <- plot(gaussSOM, what = "energy")
Let.nrg <- plot(letSOM, what = "energy")
plot(Let.nrg$data$Steps,Let.nrg$data$Energy, log="y", type="l", col=1)
lines(Gauss.nrg$data$Steps,Gauss.nrg$data$Energy,col=2)
legend("bottomleft",legend=c("Letremy","Gaussian"), lty =  c(1,1),col=c(1,2))
```

```{r}
#table(testSOM$clustering)
plot(gaussSOM, what="obs", type="hitmap")
plot(gaussSOM, what = "prototypes", type = "poly.dist", show.names = TRUE)
```

```{r}
#table(testSOM$clustering)
plot(letSOM, what="obs", type="hitmap")
plot(letSOM, what = "prototypes", type = "poly.dist", show.names = TRUE)
```

```{r}
plot(gaussSOM, what = "prototypes", type = "poly.dist", show.names = FALSE)
plot(gaussSOM, what = "prototypes", type = "umatrix")
plot(gaussSOM, what = "prototypes", type = "smooth.dist")
plot(gaussSOM, what = "prototypes", type = "mds")
#plot(testSOM, what = "prototypes", type = "grid.dist")
```
```{r}
plot(letSOM, what = "prototypes", type = "poly.dist", show.names = FALSE)
plot(letSOM, what = "prototypes", type = "umatrix")
plot(letSOM, what = "prototypes", type = "smooth.dist")
plot(letSOM, what = "prototypes", type = "mds")
#plot(testSOM, what = "prototypes", type = "grid.dist")
```

  - `"poly.dist"` represents the distances between neighboring prototypes with polygons plotted for each cell of the grid. The smaller the distance between a polygon’s vertex and a cell border, the closer the pair of prototypes. The colors indicates the number of observations in the neuron (white is used for empty neurons);

  - `"umatrix"` fills the neurons of the grid using colors that represent the average distance between the current prototype and its neighbors;

  - `"smooth.dist"` plots the mean distance between the current prototype and its neighbors with a color gradation;

  - `mds` plots the number of the neuron on a map according to a Multi Dimensional Scaling (MDS) projection;

  - `grid.dist` plots a point for each pair of prototypes, with x coordinates representing the distance between the prototypes in the input space, and y coordinates representing the distance between the corresponding neurons on the grid.

Letremy SOM seems to give more interesting results compared to Gaussian SOM.

```{r}
letClust <- letSOM$clustering
gaussClust <- gaussSOM$clustering
Letremy.Gauss.ContTable <- prop.table(table(letClust,gaussClust), margin = 2)
adj.rand.index(letClust,gaussClust)
```

```{r Letremy vs Gaussian Contingency Table}
heatmap(Letremy.Gauss.ContTable,Colv = NA, Rowv = NA, xlab = "Gaussian", ylab = "Letremy")
heatmap(Letremy.Gauss.ContTable, xlab = "Gaussian", ylab = "Letremy")
```


### Topology : square VS hexagonal

```{r}
hex.SOM <- trainSOM(X , dimension = c(10,10), topo = "hexagonal",
                    affectation = "standard", nb.save = 30, verbose = TRUE,
                    maxit = 1500, radius.type = "letremy")
save(hex.SOM, file = 'hexagonalSOM.Rdata')
```
```{r}
hex.nrg <- plot(hex.SOM, what = "energy")
plot(Let.nrg$data$Steps,Let.nrg$data$Energy, log="y", type="l", col=1)
lines(hex.nrg$data$Steps,hex.nrg$data$Energy,col=2)
legend("topright", legend=c("square","hexagonal"), lty =  c(1,1), col=c(1,2))
```

```{r}
#table(testSOM$clustering)
plot(hex.SOM, what="obs", type="hitmap")
plot(hex.SOM, what = "prototypes", type = "poly.dist", show.names = TRUE)
```

```{r}
#table(testSOM$clustering)
plot(letSOM, what="obs", type="hitmap")
plot(letSOM, what = "prototypes", type = "poly.dist", show.names = TRUE)
```

```{r}
plot(hex.SOM, what = "prototypes", type = "poly.dist", show.names = FALSE)
plot(hex.SOM, what = "prototypes", type = "umatrix")
plot(hex.SOM, what = "prototypes", type = "smooth.dist")
plot(hex.SOM, what = "prototypes", type = "mds")
#plot(testSOM, what = "prototypes", type = "grid.dist")
```
```{r}
plot(letSOM, what = "prototypes", type = "poly.dist", show.names = FALSE)
plot(letSOM, what = "prototypes", type = "umatrix")
plot(letSOM, what = "prototypes", type = "smooth.dist")
plot(letSOM, what = "prototypes", type = "mds")
#plot(testSOM, what = "prototypes", type = "grid.dist")
```

```{r}
#letClust <- letSOM$clustering
hexClust <- hex.SOM$clustering
Square.Hexa.ContTable <- prop.table(table(letClust,hexClust), margin = 2)
adj.rand.index(letClust,hexClust)
```
```{r Square vs Hexagonal Contingency Table}
heatmap(Square.Hexa.ContTable,Colv = NA, Rowv = NA, xlab = "Hexagonal", ylab = "Square")
heatmap(Square.Hexa.ContTable, xlab = "Hexagonal", ylab = "Square")
```


### Affectation : standard VS Heskes

```{r}
heskes.SOM <- trainSOM(X , dimension = c(10,10), topo = "square",
                    affectation = "heskes", nb.save = 30, verbose = TRUE,
                    maxit = 1500, radius.type = "letremy")
save(heskes.SOM, file = 'heskesSOM.Rdata')
```

```{r}
heskes.nrg <- plot(heskes.SOM, what = "energy")
plot(Let.nrg$data$Steps,Let.nrg$data$Energy, log="y", type="l", col=1)
lines(heskes.nrg$data$Steps,heskes.nrg$data$Energy,col=2)
legend("topright", legend=c("standard","heskes"), lty =  c(1,1), col=c(1,2))
```

```{r}
#table(testSOM$clustering)
plot(heskes.SOM, what="obs", type="hitmap")
plot(heskes.SOM, what = "prototypes", type = "poly.dist", show.names = TRUE)
```

```{r}
#table(testSOM$clustering)
plot(letSOM, what="obs", type="hitmap")
plot(letSOM, what = "prototypes", type = "poly.dist", show.names = TRUE)
```

```{r}
plot(heskes.SOM, what = "prototypes", type = "poly.dist", show.names = FALSE)
plot(heskes.SOM, what = "prototypes", type = "umatrix")
plot(heskes.SOM, what = "prototypes", type = "smooth.dist")
plot(heskes.SOM, what = "prototypes", type = "mds")
#plot(testSOM, what = "prototypes", type = "grid.dist")
```
```{r}
plot(letSOM, what = "prototypes", type = "poly.dist", show.names = FALSE)
plot(letSOM, what = "prototypes", type = "umatrix")
plot(letSOM, what = "prototypes", type = "smooth.dist")
plot(letSOM, what = "prototypes", type = "mds")
#plot(testSOM, what = "prototypes", type = "grid.dist")
```










```{r}
testSC <- superClass(heskes.SOM, k=20)
plot(testSC, plot.var = FALSE, show.names = FALSE)
plot(testSC, type = "grid", show.names = FALSE)
plot(testSC, what = "prototypes", type = "poly.dist", show.names = FALSE)
```


```{r}
#letClust <- letSOM$clustering
heskClust <- heskes.SOM$clustering
Stand.Heskes.ContTable <- prop.table(table(letClust,heskClust), margin = 2)
adj.rand.index(letClust,heskClust)
```

```{r Stand vs Heskes Contingency Table}
heatmap(Stand.Heskes.ContTable,Colv = NA, Rowv = NA, xlab = "heskes", ylab = "standard")
heatmap(Stand.Heskes.ContTable, xlab = "heskes", ylab = "standard")
```


## SOMbrero on Gene2vec embedding
### Gene2vec embedding result loading 

```{r}
# The following setting is important, do not omit.
options(stringsAsFactors = FALSE)
gene2vec <-  as.matrix(read.csv("./gene2vec_dim_200_iter_9.csv", header = T, row.names = 1))

dim(gene2vec)
colnames(gene2vec) <- 1:dim(gene2vec)[2] 
gene2vec[1:10,1:10]
typeof(gene2vec)
```
### Basic SOM on Gene2vec
```{r g2v default and let SOM}
g2v.default.SOM <- trainSOM(x.data = gene2vec, nb.save = 30, verbose = TRUE, maxit = 1500)
g2v.let.SOM <- trainSOM(x.data = gene2vec, nb.save = 30, verbose = TRUE, maxit = 1500,
                        radius.type = "letremy")
```
```{r Energy plot default and Letremy SOM}
g2v.default.nrg <- plot(g2v.default.SOM, what = "energy")
g2v.let.nrg <- plot(g2v.let.SOM, what = "energy")
plot(g2v.let.nrg$data$Steps,g2v.let.nrg$data$Energy, log="y", type="l", col=1)
lines(g2v.default.nrg$data$Steps,g2v.default.nrg$data$Energy,col=2)
legend("topright", legend=c("Letremy","Default"), lty =  c(1,1), col=c(1,2))
```

```{r}
plot(g2v.default.SOM, what="obs", type="hitmap")
plot(g2v.let.SOM, what="obs", type="hitmap")
plot(g2v.default.SOM, what = "prototypes", type = "poly.dist", show.names = TRUE)
plot(g2v.let.SOM, what = "prototypes", type = "poly.dist", show.names = TRUE)
```
```{r}
plot(g2v.default.SOM, what = "prototypes", type = "smooth.dist")
plot(g2v.let.SOM, what = "prototypes", type = "smooth.dist")
plot(g2v.default.SOM, what = "prototypes", type = "mds")
plot(g2v.let.SOM, what = "prototypes", type = "mds")
plot(g2v.default.SOM, what = "prototypes", type = "grid.dist")
plot(g2v.let.SOM, what = "prototypes", type = "grid.dist")
```

Once again, Letremy radius seems more interesting compared with Gaussian radius.

### Load t-SNE representation of Gene2vec data
```{r load t-SNE for gene2vec}
g2v.tSNE <- read.csv(file = "gene2vec_TSNE.csv", header = FALSE)
head(g2v.tSNE)
```

### Clusters representation on t-SNE projection

```{r plotClustering function}
plotClustering <- function(proj2D, SOMres) {
  N <- dim(SOMres$prototypes)[1] 
  
  mds <- plot(SOMres, what = "prototypes", type = "mds")$data
  scalex <- (mds$x - min(mds$x))/(max(mds$x)-min(mds$x))
  scaley <- (mds$y - min(mds$y))/(max(mds$y)-min(mds$y))
  colClust <- rgb(scalex,scaley,0.5) 

  colors <- colClust[SOMres$clustering] 
  #ClustOrder <- superClass(SOMres)$tree$order
  #colors <- (paletteer_c("pals::jet", n = N))[match(1:N,ClustOrder)]
  
  #kovesi.diverging_rainbow_bgymr_45_85_c67 
  plotCol <- colors[ unclass(SOMres$clustering) ]
  plot(proj2D, pch = 20, cex = 0.3, col = plotCol)
}
```

```{r}
plotClustering(g2v.tSNE, g2v.default.SOM)
plotClustering(g2v.tSNE, g2v.let.SOM)
```
Similar clusters share similar colors (+/-)

### Test all parameters 
```{r}
dimSOM <- c(10,10)
maxitSOM <- 1000
nb.saveSOM <- 10

topoSOM <- c("square", "hexagonal")
radiusSOM <- c("gaussian", "letremy")
distSOM <- c("letremy", "euclidean", "maximum", "manhattan", "canberra", "minkowski")
affectSOM <- c("standard", "heskes")

names <- list.files("./SOM/")
```

```{r}
for (t in topoSOM) {
  for (r in radiusSOM) {
    
    if (t == "hexagonal") {
      distSOM <- c("euclidean")
    } else if (r == "gaussian") {
      distSOM <- c("euclidean", "maximum", "manhattan", "canberra", "minkowski")
    } else {
      distSOM <- c("letremy", "euclidean", "maximum", "manhattan", "canberra", "minkowski")
    }
    
    for (d in distSOM) {
      for (a in affectSOM) {
        name <- paste("som_",t,"_",r,"_",d,"_",a,".RData", sep = "")
        if (!(name %in% names)){
          som <- try(trainSOM(x.data = gene2vec, dimension = dimSOM, topo = t, 
                        radius.type = r, dist.type = d, affectation = a,
                        maxit = maxitSOM, nb.save = nb.saveSOM, verbose = TRUE))
          if (inherits(som, "try-error")) {
            next
          }
          name2 <- paste("./SOM/",name, sep = "")
          save(som, file = name)
          names <- append(names, name)
        } 
      }
    }
  }
}
```
```{r}
names <- list.files("./SOM/")
for (name in names) {
  f = paste("./SOM/",name,sep = "")
  load(f)
  plotProto(som)
  title(name)
  plotClustering(g2v.tSNE, som)
  title(name)
}

```
```{r plotProto function}
plotProto <- function(som){
  mds <- plot(som, what = "prototypes", type = "mds")$data
  
  scalex <- (mds$x - min(mds$x))/(max(mds$x)-min(mds$x))
  scaley <- (mds$y - min(mds$y))/(max(mds$y)-min(mds$y))
  
  colClust <- rgb(scalex,scaley,0.5) 
  
  plot(scalex,scaley,pch=20,col=colClust,cex=2*table(som$clustering)/max(table(som$clustering)))  
} 

```


### Clusters functionnal enrichment analysis
```{r Setup Enrichr}
setEnrichrSite("Enrichr") # Human genes
websiteLive <- TRUE
dbs <- listEnrichrDbs()
if (is.null(dbs)) websiteLive <- FALSE
if (websiteLive) head(dbs)
dbs[grep('GO', dbs$libraryName),c(1,3)] 
```

```{r enrichment analysis}
dbs <- c("GO_Molecular_Function_2021", "GO_Cellular_Component_2021", "GO_Biological_Process_2021")
absPath <- '~/Documents/data/SOM/'
SOMfiles <- list.files(absPath)[grep('som_', list.files(absPath))] 
clusters <- 1:dim(som$prototypes)[1]
namesID <- sub(pattern = ".*som","",SOMfiles)
enrNames <- paste("enr",namesID,sep="")

for (i in 1:length(SOMfiles)) {
  f <- paste(absPath,SOMfiles[i],sep = "")
  load(f)
  enrClust <- list()
  for (clust in clusters) {
     enriched <- enrichr(names(som$clustering[
       som$clustering == clust]), dbs)
     clusterName <- paste("cl",clust,sep = "")
     enrClust[[clusterName]] <- enriched
  }
  fenr <- paste(absPath,enrNames[i],sep = "")
  save(enrClust, file = fenr)
}
  

```

```{r print result}
clNames <- names(enrClust)
for (cl in clNames) {
#  plt <- plotEnrich(enrClust[[cl]][[3]], showTerms = 20, numChar = 40, y = "Count", orderBy = "P.value")
#  print(plt)
  print(head(enrClust[[cl]][[3]][,c(1,4)]))
}
```


We will count the number of clusters with at least one adjusted p-value <= 0.05
```{r count enriched clusters}
absPath <- '~/Documents/data/SOM/'
enrfiles <- list.files(absPath)[grep('enr_', list.files(absPath))] 

signSOM <- rep(0,length(enrfiles))
for (i in 1:length(enrfiles)){
  #print(i)
  f <- paste(absPath,enrfiles[i], sep = "")
  load(f)
  clNames <- names(enrClust)
  #print(clNames)
  for (cl in clNames) {
    #print(cl)
    maxpval <- enrClust[[cl]][[3]][1,4]
    #print(maxpval)
    if (!is.na(maxpval) && maxpval<=0.05) signSOM[i] <- signSOM[i]+1  
  }
} 
as.data.frame(signSOM, row.names = enrfiles)
```

There is a clear gap between SOM using standard affectation and Heskes affectation. Letremy seems a bit better but a lot of the parameter combination didn't work so it is hard to compare...


```{r}

load('~/Documents/data/SOM/som_square_gaussian_euclidean_standard.RData')
g2v.dist <- dist(gene2vec, method = "euclidean")
save(g2v.dist, file = '~/Documents/data/g2v_eucDist.RData')
g2v.dist.mat <- as.matrix(g2v.dist)
sil <- silhouette(som$clustering, g2v.dist.mat)
summary(sil)

sil2 <- sil[order(sil[,1], sil[,3]),]
tab2 <- table(sil2[,1])
N = length(tab2)
colors <- sample(paletteer_c("pals::jet", n = N),N)
col2 <- rep(colors, tab2)
barplot(sil2[,3], col = col2[], space = 0, border=NA)
```
A lot of silhouette coeff are below 0 -> :(

```{r}
dendsom <- superClass(som)
summary(dendsom$tree)
head(dendsom$tree$height)
cut25  <- 0.01*max(dendsom$tree$height)
cutbool <- dendsom$tree$height > cut25
k <- sum(cutbool)+1

my.sc <- superClass(som, k = k)
summary(my.sc)
plot(my.sc, plot.var = FALSE)
recluster <- my.sc$cluster[som$clustering] 
names(recluster) <- names(som$clustering)

#g2v.dist.mat
```

```{r}
plotSil <- function(clust, distc){
  sil <- silhouette(clust, distc)

  sil2 <- sil[order(sil[,1], sil[,3]),]
  tab2 <- table(sil2[,1])
  N = length(tab2)
  colors <- sample(paletteer_c("pals::jet", n = N),N)
  col2 <- rep(colors, tab2)
  barplot(sil2[,3], col = col2[], space = 0, border=NA)
} 

```

```{r}
plotSil(som$clustering, g2v.dist.mat)
plotSil(recluster, g2v.dist.mat)
```
...not that better...

```{r}
absPath <- '~/Documents/data/SOM/'
somfiles <- list.files(absPath)[grep('som_', list.files(absPath))]
somfiles <- somfiles[-grep('heskes',somfiles)] 

for (name in somfiles) {
  f = paste(absPath,name,sep = "")
  load(f)
  plotSil(som$clustering, g2v.dist.mat)
  title(name)
}

```
```{r}
distonce = c("minkowski")
for (d in distonce) {
  g2v.dist <- dist(gene2vec, method = d)
  g2v.dist.mat <- as.matrix(g2v.dist)
  save(g2v.dist.mat, file = paste('~/Documents/data/g2v_',d,'Dist.RData', 
                                  sep = ""))
}
```

```{r}
absPath <- '~/Documents/data/SOM/'
somfiles <- list.files(absPath)[grep('som_', list.files(absPath))]
somfiles <- somfiles[-grep('heskes',somfiles)] 

for (name in somfiles) {
  f <- paste(absPath,name,sep = "")
  load(f)
  d <- strsplit(name, "_")[[1]][4] 
  if (d=="letremy") d <- "euclidean"
  load(paste('~/Documents/data/g2v_',d,'Dist.RData', 
                                  sep = ""))
  plotSil(som$clustering, g2v.dist.mat)
  title(name)
}
```
Letremy seems better

### Hexagonal topology SOM

```{r Hexagonal topology SOM}
g2v.hex.SOM <- trainSOM(x.data = gene2vec, nb.save = 30, verbose = TRUE, maxit = 1500,
                        radius.type = "gaussian", topo = "hexagonal")
g2v.hexlet.SOM <- trainSOM(x.data = gene2vec, nb.save = 30, verbose = TRUE, maxit = 1500,
                        radius.type = "letremy", topo = "hexagonal")
name <- paste(absPath,"som_hexagonal_letremy_euclidean_standard.RData", sep = "")
som <- g2v.hexlet.SOM
save(som, file = name)
```

```{r Plot hex clustering}
plotProto(g2v.hex.SOM)
title("hexagonal gaussian")
plotClustering(g2v.tSNE, g2v.hex.SOM)
title("hexagonal gaussian")

plotProto(g2v.hexlet.SOM)
title("hexagonal letremy")
plotClustering(g2v.tSNE, g2v.hexlet.SOM)
title("hexagonal metremy")

```
```{r Hexagonal topo enrichment analysis}
dbs <- c("GO_Molecular_Function_2021", "GO_Cellular_Component_2021", "GO_Biological_Process_2021")

somhex <- list(g2v.hex.SOM, g2v.hexlet.SOM)
nameshex <- c("som_hexagonal_gaussian_euclidean_standard.RData","som_hexagonal_letremy_euclidean_standard.RData")
namesID <- sub(pattern = ".*som","",nameshex)
enrNames <- paste("enr",namesID,sep="")
for (i in 1:length(somhex)) {
  som <- somhex[[i]] 
  clusters <- 1:dim(som$prototypes)[1]
  enrClust <- list()
  for (clust in clusters) {
     enriched <- enrichr(names(som$clustering[
       som$clustering == clust]), dbs)
     clusterName <- paste("cl",clust,sep = "")
     enrClust[[clusterName]] <- enriched
  }
  fenr <- paste(absPath,enrNames[i],sep = "")
  save(enrClust, file = fenr)
}
  
```

```{r}
plotSil(g2v.hex.SOM, g2v.dist.mat)
plotSil(g2v.hexlet.SOM, g2v.dist.mat)
```
Hexagonal topology with Letremy radius seems to be the better result we have seen so far !

```{r}
plot(superClass(g2v.hexlet.SOM))
```



```{r}
som <- trainSOM(x.data = gene2vec, nb.save = 30, verbose = TRUE, maxit = 1500,
                        radius.type = "letremy", topo = "hexagonal",dimension = c(7, 7))
```

```{r}
plot(som, what="obs", type="hitmap")
plot(som, what = "prototypes", type = "poly.dist", show.names = TRUE)
plot(som, what = "prototypes", type = "smooth.dist")
plot(som, what = "prototypes", type = "mds")
plot(som, what = "prototypes", type = "grid.dist")
```

```{r}
quality(som)
```
The topographic error value varies between 0 (good projection quality) and 1 (poor projection quality). Here, the topographic quality of the mapping is equal to 0.50, which means that around 50% of the observations have a second best unit in the neighborhood of the best matching unit.

The quantization error is an unbounded positive number. The closer it is to 0, the better the projection quality.


```{r}
ndim <- 2:30
topo <- c()
quan <- c()
for (n in ndim) {
  som <- trainSOM(x.data = gene2vec, verbose = FALSE, maxit = 1000, 
                  radius.type = "letremy", topo = "hexagonal", 
                  dimension = c(n, n), dist.type = "euclidean")
  q <- quality(som)
  topo <- append(topo, q$topographic)
  quan <- append(quan, q$quantization)
}

plot(ndim,quan)
plot(ndim,topo)

save(ndim, quan, topo, file = paste(absPath, "qualityVSndim.RData"))
```

```{r}
ndim <- 5:25
nrep <- 10
topoMean <- c()
quanMean <- c()
topoMed <- c()
quanMed <- c()
for (n in ndim) {
  to <- c()
  qu <- c()
  for(i in 1:nrep){
    som <- trainSOM(x.data = gene2vec, verbose = FALSE, maxit = 1000, 
                    radius.type = "letremy", topo = "hexagonal", 
                    dimension = c(n, n), dist.type = "euclidean")
    q <- quality(som)
    to <- append(to, q$topographic)
    qu <- append(qu, q$quantization)
  }
  topoMean <- append(topoMean, mean(to))
  quanMean <- append(quanMean, mean(qu))
  topoMed <- append(topoMed, median(to))
  quanMed <- append(quanMed, median(qu))
}

#plot(ndim,quan)
#plot(ndim,topo)

save(ndim, topoMean, quanMean, topoMed, quanMed, file = paste(absPath, "qualityMeanMedVSndim.RData"))
```

## Create membership matrix

```{r data sets loading}
options(stringsAsFactors = FALSE)
#rna <- t(read.csv("../data/DataRnaseq", header = T, sep = " "))
rna <- read.csv("../data/DataRnaseq_Transposed.csv", header = T, row.names=1)
g2v <- read.csv("../data/gene2vec_dim_200_iter_9.csv", header = T, row.names=1)
```


```{r SOM}
n <- 15
SOM_rna <- trainSOM(x.data = rna, dimension = c(n,n), topo = "hexagonal", radius.type = "letremy", dist.type = "euclidean", affectation = "standard", maxit = 1000)
SOM_g2v <- trainSOM(x.data = g2v, dimension = c(n,n), topo = "hexagonal", radius.type = "gaussian", dist.type = "euclidean", affectation = "standard")
```

```{r tSNE loading}
g2v_tSNE <- read.csv(file = "../data/gene2vec_TSNE.csv", header = FALSE)
rna_tSNE <- read.csv(file = "../data/RNAseq_tSNE.csv", header = TRUE, row.names = 1)
```


```{r plot RNA}
plotProto(SOM_rna)
title("RNA-seq SOM prototypes")
plotClustering(rna_tSNE, SOM_rna)
title("RNA-seq SOM clustering")
```
```{r plot Gene2vec}
plotProto(SOM_g2v)
title("Gene2vec SOM prototypes")
plotClustering(g2v_tSNE, SOM_g2v)
title("Gene2vec SOM clustering")
```
```{r}
createMM <- function(SOM_, filename = FALSE){
  genes <- names(SOM_$clustering)
  cl <- sort(unique(SOM_$clustering))
  nclust <- length(cl)
  mm <- data.frame(matrix(0, nrow = length(genes), ncol = nclust), row.names = genes)
  colnames(mm) <- cl
  #print(mm)
  for (i in 1:length(genes)) {
    #print(genes[i])
    #print(SOM_$clustering[genes[i]])
    mm[genes[i],as.character(SOM_$clustering[genes[i]])] <- 1 
  }
  if (filename != FALSE) {
    write.csv(mm, file = paste(filename, "csv", sep = "."), row.names = TRUE)
  }
  mm
}
```
```{r}
rna_mm <- createMM(SOM_rna, filename = "./ModuleMatrix/SOM_rna_HexLet")
g2v_mm <- createMM(SOM_g2v, filename = "./ModuleMatrix/SOM_g2v_HexGauss")
```
```{r}
dim(rna_mm)
dim(g2v_mm)
```







# Create membership matrices
```{r}
setwd('~/Documents/Clustering/')
source('./clusterian_R/lab2mat.R')

# The following setting is important, do not omit.
options(stringsAsFactors = FALSE)
g2v <-  as.matrix(read.csv("../data/gene2vec_dim_200_iter_9.csv", header = T, row.names = 1))
head(row.names(g2v))
```

```{r}
load('./SOM/som_hexagonal_gaussian_euclidean_standard.RData')
mm <- lab2mat(som$clustering, genes = row.names(g2v))
head(mm)
colSums(mm)
```


# Re-test all parameters

All parameters are re-tested with a $15\times 15$ grid.

```{r}
library(SOMbrero)
```

```{r}
options(stringsAsFactors = FALSE)
rna <- read.csv("../data/DataRnaseq_Transposed.csv", header = T, row.names = 1)
```

```{r}
dimSOM <- c(15,15)
maxitSOM <- 1000
nb.saveSOM <- 5

topoSOM <- c("square", "hexagonal")
radiusSOM <- c("gaussian", "letremy")
distSOM <- c("letremy", "euclidean", "maximum", "manhattan", "canberra", "minkowski")
# affectSOM <- c("standard", "heskes")
affectSOM <- c("standard")


```

```{r}
names <- list.files("./SOM/")
for (t in topoSOM) {
  for (r in radiusSOM) {
    
    if (t == "hexagonal") {
      distSOM <- c("euclidean")
    } else if (r == "gaussian") {
      distSOM <- c("euclidean", "maximum", "manhattan", "canberra", "minkowski")
    } else {
      distSOM <- c("letremy", "euclidean", "maximum", "manhattan", "canberra", "minkowski")
    }
    
    for (d in distSOM) {
      for (a in affectSOM) {
        name <- paste("SOM_rna_",t,"_",r,"_",d,"_",a,".RData", sep = "")
        if (!(name %in% names)){
          som <- try(trainSOM(x.data = rna, dimension = dimSOM, topo = t, 
                        radius.type = r, dist.type = d, affectation = a,
                        maxit = maxitSOM, nb.save = nb.saveSOM, verbose = TRUE))
          if (inherits(som, "try-error")) {
            next
          }
          name2 <- paste("./SOM/",name, sep = "")
          save(som, file = name2)
          names <- append(names, name)
        } 
      }
    }
  }
}
```

```{r}
source('~/Documents/Clustering/clusterian_R/lab2mat.R')
directory <- "./SOM/"
names <- list.files(directory)
mm_dir <- paste(directory,'MembMatrix/', sep='')
mm_names <- list.files(mm_dir)
for (file in names) {
  file_split <- strsplit(file, "\\.")[[1]]
  name <- file_split[-length(file_split)]
  ext <- file_split[length(file_split)]
  mm_file <- paste(name,'.csv',sep="")
  #print(ext)
  if (ext=='RData' && !(mm_file %in% mm_dir)) {
    path_ <- paste(directory,file, sep = "")
    load(path_)
    mm <- lab2mat(som$clustering, genes = names(som$clustering))
    write.csv(mm,paste(mm_dir, mm_file, sep = ""), row.names = TRUE)
  }
}
```

